%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  New template code for TAMU Theses and Dissertations starting Spring 2021.
%
%
%  Author: Thesis Office
%
%  Last Updated: 1/13/2021
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                           ABSTRACT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter*{ABSTRACT}
\addcontentsline{toc}{chapter}{ABSTRACT} % Needs to be set to part, so the TOC doesn't add 'CHAPTER ' prefix in the TOC.

\pagestyle{plain} % No headers, just page numbers
\pagenumbering{roman} % Roman numerals
\setcounter{page}{2}


Many combinatorial optimization (CO) problems are naturally defined on graphs, arising
in domains such as social networks, transportation, and telecommunications. Because
these problems are typically NP-hard, even the best heuristics struggle as real-world
graphs grow to millions or billions of nodes. To scale these solvers, a natural strategy is to prune the search
space by discarding elements unlikely to contribute to a high-quality solution, so that
solvers operate on a much smaller instance. This research develops a unified approach to
search space pruning for graph CO that spans classical approaches with theoretical guarantees, learning-based
methods, and large language model-driven automation.

We begin with the classical foundations of pruning and introduce QuickPrune, a
lightweight, single-pass algorithm that provides provable guarantees on both the fraction
of optimal value retained and the size of the resulting pruned ground set under weak
submodularity assumptions.

To overcome the sequential nature of classical approaches and the domain expertise
required in algorithm design, we propose Hierarchical DeepPruner, an adaptive two-stage
framework that combines graph neural network pre-pruning with neural Monte Carlo Tree
Search fine-pruning, eliminating the need for domain-specific feature engineering while
achieving competitive performance across diverse graph CO problems under size
constraints, even when trained on small synthetic graphs.

Building on this foundation, we introduce LLM2Prune, a framework that leverages large
language models as domain experts to automatically discover problem-specific features through an LLM-guided beam search, enabling scalable and constraint-adaptive
pruning without human intervention.

Taken together, this work presents a unified view of search space reduction for graph CO,
from provable classical algorithms to problem-agnostic deep learning and automated
feature discovery, demonstrating that efficient pruning can preserve near-optimal solution
quality while reducing ground sets by over 90\%, enabling optimization at scales
previously intractable.

\pagebreak{}
