\pagestyle{plain} % No headers, just page numbers

\chapter{\uppercase {Related Work}}

The proposed work draws on three areas of prior research: classical pruning methods,
learning-based pruning methods, and automated feature engineering using large language
models. This chapter reviews existing work in each area and highlights the limitations
that motivate the contributions of this proposal.

\section{Classical Approaches to Pruning}
Classical pruning methods for combinatorial optimization (CO) fall into two categories:
problem-specific heuristics and general-purpose algorithms. Problem-specific heuristics
exploit the structure of a single CO problem to discard unpromising elements, such as
degree-based candidate filtering for influence
maximization~\cite{chen2010scalable} and edge elimination rules for the traveling
salesman problem~\cite{sun2021generalization}. While effective within their respective
domains, these techniques do not transfer to other CO problems.

On the general-purpose side, \cite{zhou2017scaling} propose a pruning algorithm grounded
in submodularity to reduce the cost of submodular maximization under size constraints.
Their method constructs a pruned submodularity graph to identify and discard elements with
low marginal contributions. However, this approach processes elements sequentially, which
limits its scalability to large instances. More broadly, all existing classical pruning
methods share three limitations: they target a single instance of size-constrained
maximization, they provide no mechanism to leverage patterns across problem instances, and
they lack formal guarantees on the quality of the pruned ground set.

\section{Learning-Based Approaches to Pruning}
Learning-based methods take a data-driven approach to pruning, training models on solved
instances to identify elements that can be removed from the ground set.
These methods can be broadly organized into three categories: supervised approaches with
hand-crafted features, GNN-based approaches, and hybrid methods combining supervised and
reinforcement learning.

\textbf{Supervised approaches with hand-crafted features.}
The earliest learning-based pruning methods train binary classifiers on small, pre-solved
instances to predict whether a vertex or edge belongs to an optimal solution. For Maximum
Clique Enumeration, \cite{lauri2019fine} apply logistic regression with hand-crafted
graph-theoretic features, and \cite{grassia2019learning, lauri2023learning} extend this
with multi-stage classifiers that iteratively reduce the search space. For the Traveling
Salesman Problem, \cite{fitzpatrick2021learning, sun2021generalization} train classifiers
to prune edges unlikely to belong to the optimal tour. A common limitation across these
methods is their heavy reliance on problem-specific feature engineering, which restricts
generalization to new CO problems.

\textbf{GNN-based approaches.}
To reduce the dependence on hand-crafted features, \cite{tian2024combhelper} propose
COMBHelper, a GNN-based framework with knowledge distillation and a problem-specific
boosting module designed to generalize across multiple CO problems. While COMBHelper
represents a step toward problem-agnostic pruning, it still relies on computationally
expensive Fourier-based feature transformations and a problem-specific boosting mechanism for strong
performance.

\textbf{Hybrid supervised and reinforcement learning methods.}
Hybrid methods that combine supervised learning with reinforcement learning are more
problem-agnostic. GCOMB~\cite{manchanda2020gcomb} uses a probabilistic greedy
algorithm to generate training data, then applies a weighted-degree heuristic to prune
the ground set before training a Q-learning network on the reduced search space. However,
its initial pruning depends on a handcrafted degree heuristic that does not generalize
well across CO problems~\cite{ireland2022lense}. LeNSE~\cite{ireland2022lense} takes a
different approach: a GNN is first trained to generate embeddings for sampled subgraphs,
and a policy network then uses these embeddings to guide a local search toward subgraphs
likely to contain the optimal solution. A major drawback is its reliance on domain
expertise to categorize subgraphs, set class-specific sample counts, and tune
hyperparameters such as subgraph size and embedding dimensionality separately for each
dataset and problem.

Despite these advances, no existing learning-based pruning method simultaneously achieves
problem-agnostic design, freedom from hand-crafted features, and scalability to large
graph instances. Moreover, all existing methods are designed and evaluated exclusively
under size constraints, leaving other constraint types such as knapsack constraints
unaddressed.

\section{Automated Feature Engineering via LLMs}
Automated feature engineering seeks to generate useful features from raw data without
human effort. Prior work with large language models (LLMs) has focused primarily on
tabular data, where new features are derived from existing columns. These approaches
either rely on manually designed feature-generation
rules~\cite{zhang2023openfe, zhang2024dynamic} or few-shot
prompts~\cite{han2024large}, and thus still depend on hand-crafted design choices.
Moreover, existing methods~\cite{nam2024optimized, hollmann2023large} suffer from error
propagation caused by simple feedback loops. These approaches are distinct from methods
that directly generate algorithms using LLMs~\cite{yang2023large, liu2024evolution},
which often exhibit degraded performance as graph size increases.

LLM-driven feature engineering is a natural fit for pruning, as LLMs can encode the kind
of domain knowledge that would otherwise require expert design of problem-specific
features. To the best of our knowledge, no prior work has applied LLM-driven feature
engineering to search space reduction for graph CO problems.